---
title: "analysis"
author: "Rishabh Verma"
date: "7/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
```

```{r}
data <- read_csv("data/clock.csv") %>% arrange(time)
```

Type -1: uptimeMillis (for touch events)
Type -2: elapsedRealtimeNanos (for sensor events)
Type -3: delta (for converting)

The first tidbits of types -1,-2 will be the ones used to compute the sole -3.

```{r}
first_millis <- data %>% filter(type == -1) %>% head(1) %>% select(time) %>% as.double()

first_nanos <- data %>% filter(type == -2) %>% head(1) %>% select(time) %>% as.double()

delta <- data %>% filter(type == -3) %>% select(time) %>% as.double()
delta

```

This represents the error computed by the delta function in unit nanoseconds. It's only 2.5 millliseconds off, which is reasonable.
```{r}
first_millis*1000 + delta - first_nanos
```

```{r}
millis <- data %>% filter(type == -1)
nanos <- data %>% filter(type == -2)

millis
nanos
```

There are four rows in each table corresponding to 4 synchronized clock measurements. One ring to bring them all, and in the darkness, bind them.


```{r}
events <- bind_cols(millis, nanos) %>% select(type_1 = type...1, time_1 = time...2, type_2 = type...6, time_2 = time...7)

head(events)

events %>% ggplot(aes(x=time_1, y=time_2)) +
  geom_point() +
  geom_line()
```
This does look linear. Let's compute the linear regression.

```{r}
x <- events$time_1 %>% as.data.frame()
y <- events$time_2 %>% as.data.frame()


relation <- lm(time_1 ~ time_2, data = events)

print(relation)
```



Type 1: accelerometer
Type 4: gyroscope
Type 6: pressure
Type -27: touch


```{r}
accelerometer <- data %>% 
  filter(type == 1) %>% 
  select(time, one, two, three) %>% 
  arrange(desc(time))
  
pressure <- data %>% 
  filter(type == 6) %>% 
  select(time, one) %>% 
  arrange(desc(time))

gyroscope <- data %>% 
  filter(type == 4) %>% 
  select(time, one, two, three) %>% 
  arrange(desc(time))

touch <- data %>% 
  filter(type == -27) %>% 
  select(time, one, two) %>% 
  arrange(desc(time))
```




```{r}
accelerometer %>% select(time) %>% mutate(time = time) %>% summary()

gyroscope %>% select(time) %>% mutate(time = time) %>% summary()

pressure %>% select(time) %>% mutate(time = time) %>% summary()

touch %>% select(time) %>% mutate(time = time %% 1e+10) %>% summary()
```

```{r}
first %% 1e+11
last %% 1e+11
cutoff %% 1e+11
```

```{r}
p = 1.00

first <- data %>% filter(type == -27) %>% select(time) %>% min()
last  <- data %>% filter(type == -27) %>% select(time) %>% max()

cutoff = (last-first)*p + first
```

Graph the first `p`*100 percent of data.
```{r}
accelerometer %>% 
  filter(time < cutoff) %>% 
  ggplot() +
  geom_point(aes(x=time,y=one))

pressure %>% 
  filter(time < cutoff) %>% 
  ggplot() +
  geom_point(aes(x=time,y=one))

gyroscope %>% 
  filter(time < cutoff) %>% 
  ggplot() +
  geom_point(aes(x=time,y=one))

```

Graph all the data.
```{r}
accelerometer %>% 
  ggplot() +
  geom_point(aes(x=time,y=one))

pressure %>% 
  ggplot() +
  geom_point(aes(x=time,y=one))

gyroscope %>% 
  ggplot() +
  geom_point(aes(x=time,y=one))

touch %>%
  ggplot() +
  geom_point(aes(x=time,y=0))
```


```{r}
p = 0.03

gyroscope <- data %>% 
  filter(type == 4) %>% 
  select(time, one, two, three) %>% 
  arrange(desc(time))


first <- gyroscope %>% select(time) %>% min()
last <- gyroscope %>% select(time) %>% max()

cutoff = (last-first)*p + first

gyroscope %>% 
  filter(time < cutoff) %>% 
  ggplot() +
  geom_point(aes(x=time,y=one))

head(gyroscope)
```

