---
title: "Big Brother Barometer"
author: "Rishabh Verma"
date: "2022-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      warning=FALSE,
                      message=FALSE)

library(tidyverse)
library(ggthemes)

## Filepath on my desktop
setwd("E:\\Desktop\\projects\\docs-BBBarometer")
```

## Introduction


This project will show you how your smartphone's accelerometer and pressure sensor can be used as a side-channel attack to read your PIN. 

(scroll to continue)

## Plan of Attack

```{r load_data, include=FALSE}
tidbits <- read_csv("data/trial.csv")
source("cleaning_and_plotting.R")
data <- clean_tidbits(tidbits)
rm(tidbits)

default_range = data %>% filter(type=="touch") %>% pull(time) %>% quantile(c(0.1,0.2))
```

Let's start by typing in a bunch of random numbers.  

> video appears  

> side-scrolling graph appears with red touches


```{r}
tap_circles <- data %>%
  filter(type=="touch") %>%
  ggplot(aes(x=time, y=0)) +
  geom_point(color="red", size=5) +
  coord_fixed() +
  scale_y_continuous(position="right", limits=0.2*diff(default_range)/2*c(-1,1)) +
  scale_x_reverse(limits=rev(default_range))
tap_circles
```


Your phone's barometer is normally used to measure the air pressure.  
When you tap your phone screen, the barometer records a sharp increase in air pressure.  

> side-scrolling pressure graph appears with red touches









