---
title: "Big Brother Barometer"
author: "Rishabh Verma"
date: "2022-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      warning=FALSE,
                      message=FALSE)

library(tidyverse)
library(ggthemes)
library(ggpubr) # for multiple ggplots in one figure
library(fuzzyjoin)  # for mining a pressure feature

source("cleaning_and_plotting.R")
source("phase_one_functions.R")
source("phase_two_functions.R")

## Filepath on my desktop
setwd("E:\\Desktop\\projects\\docs-BBBarometer")
```

## Introduction


This project will show you how your smartphone's accelerometer and pressure sensor can be used as a side-channel attack to read your PIN. 



## Plan of Attack




```{r load_data, include=FALSE}
tidbits <- read_csv("data/trial.csv")

data <- clean_tidbits(tidbits)
rm(tidbits)
default_range = data %>% filter(type=="touch") %>% pull(time) %>% quantile(c(0.1,0.2))
prop = 0.5
data <- data %>%
  mutate(set = ifelse(time < quantile(time, prop),
                      "training", "testing"))
exp1 <- build_experiment(data=data)
rm(data)

attach(exp1)
# this creates:
# - data
# - touch_confidence
# - threshold
# - threshold_trials
# - events
```
To start, we need a game that will have the user tapping on different parts of the screen. 
Whack-a-mole should do the trick.


> video appears


```{r, fig.height=2}
tap_circles <- data %>%
  filter(type=="touch") %>%
  ggplot(aes(x=time, y=0)) +
  geom_point(color="red", size=5) +
  coord_fixed() +
  scale_y_continuous(position="right", limits=0.08*diff(default_range)/2*c(-1,1)) +
  scale_x_reverse(limits=rev(default_range))
tap_circles
```


A phone's barometer is normally used to measure atmospheric air pressure, but funnily enough, when a user taps their phone screen, the barometer records a sharp increase in air pressure.

```{r, fig.height=3}
pressure_plot <- scatterplot(data %>% filter(type %in% c("pressure", "touch")) %>% mutate(one = one+980), "pressure", "one") +
  scale_y_continuous(position="right", limits=c(979.9, 980.15)) +
  xlim(default_range) +
  labs(title="", subtitle="") +
  theme(legend.position="none",
        axis.title.x = element_blank())

ggarrange(pressure_plot, tap_circles,
          ncol=1)
```

At the same time, the gyroscope records the rotational motion of the phone, 
It's very sensitive, but it does pick up the impulse delivered by a user's finger.

```{r}

gyro = list()

gyro$p1 = scatterplot(data %>% filter(type %in% c("gyroscope", "touch")), "gyroscope", "one")
gyro$p2 = scatterplot(data %>% filter(type %in% c("gyroscope", "touch")), "gyroscope", "two")
gyro$p3 = scatterplot(data %>% filter(type %in% c("gyroscope", "touch")), "gyroscope", "three")

gyro$p1 = gyro$p1 + scale_y_continuous(position="right", limits=0.6*c(-1,1)) +
  ylim(0.5*c(-1,1)) +
  xlim(default_range) +
  labs(title="", subtitle="") +
  theme(legend.position="none",
        plot.margin=unit(c(0,0,0,0), 'cm'),
        axis.title.x = element_blank()) +
  scale_y_continuous(position="right")
gyro$p2 = gyro$p2 + scale_y_continuous(position="right", limits=0.6*c(-1,1)) +
  ylim(0.5*c(-1,1)) +
  xlim(default_range) +
  labs(title="", subtitle="") +
  theme(legend.position="none",
        plot.margin=unit(c(0,0,0,0), 'cm'),
        axis.title.x = element_blank()) +
  scale_y_continuous(position="right")
gyro$p3 = gyro$p3 + scale_y_continuous(position="right", limits=0.6*c(-1,1)) +
  ylim(0.5*c(-1,1)) +
  xlim(default_range) +
  labs(title="", subtitle="") +
  theme(legend.position="none",
        plot.margin=unit(c(0,0,0,0), 'cm'),
        axis.title.x = element_blank()) +
  scale_y_continuous(position="right")


ggarrange(pressure_plot, gyro$p1, gyro$p2, gyro$p3, tap_circles,
          heights=c(1,1,1,1,1),
          ncol=1)
```

If the __pressure sensor__ reveals when the user taps,  
and the _gyroscope_ then reveals where the user taps,  
then both sensors can be used to reveal a user's PIN code.

